---
title: "Pathways and metabolite sets in metpath"
author:
- name: Xiaotao Shen PhD (https://www.shenxt.info/)
date: "Created on 2020-03-28 and updated on `r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: no
  pdf_document:
    toc: no
vignette: >
  %\VignetteIndexEntry{pahtway_database}
  %\VignettePackage{metpath}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

```{r setup}
library(metpath)
```

# **Brief introduction**

Two widely used metabolic pathways are utilized in `metpath`, namely [`KEGG`](https://www.genome.jp/kegg/) and [`HMDB`/`SMPDB`](https://smpdb.ca/).

# **KEGG pathway**

We can just download the `KEGG` online or from the `metpaht` pacakge.

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
library(metpath)
```


```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
kegg_hsa_pathway = 
  get_kegg_pathway(local = TRUE, organism = "hsa", threads = 3)
```

The date when the pathway database is downloaded will show.

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
kegg_hsa_pathway
```

Just type the pathway and the information of it will show.

We can also download the latest version of `KEGG` pathway. Make sure the internet is good and it will take a while to download.

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
kegg_hsa_pathway2 = 
  get_kegg_pathway(local = FALSE, organism = "hsa", threads = 3)
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
kegg_hsa_pathway2
```

# **KEGG compound**

We can also get the KEGG compound database using `metpath`.

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
kegg_compound = 
  get_kegg_compound(local = TRUE, threads = 5)
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
kegg_compound
```

We can also download the online `KEGG` compound database. But it will take a long time.

```{r,eval=FALSE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
kegg_compound2 = 
  get_kegg_compound(local = FALSE, threads = 5)
```


# **HMDB pathway**

We can just download the `HMDB` online or from the `metpaht` pacakge.

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
hmdb_hsa_pathway = 
  get_hmdb_pathway(threads = 3)
```

The date when the pathway database is downloaded will show.

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
hmdb_hsa_pathway
```

Just type the pathway and the information of it will show.


# **HMDB compound**

We can also get the HMDB compound database using `metpath`.

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
hmdb_compound = 
  get_hmdb_compound(threads = 5)
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
hmdb_compound
```


# **Processing of pathway class**

## Brief information of pathway database

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
length(kegg_hsa_pathway)
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
get_pathway_class(kegg_hsa_pathway)
```

## Accessor

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
kegg_hsa_pathway[1:5]
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
head(names(kegg_hsa_pathway))
```

```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
names(kegg_hsa_pathway[c("hsa00010", "hsa00020")])
```


## Extract data


```{r,eval=TRUE,warning=FALSE, R.options="", message=FALSE, cache=TRUE}
database_info(kegg_hsa_pathway)
```